\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{geometry}
\geometry{margin=1in}

\begin{document}

\section*{The Massive Function}

The single formulation is given by:

\[
\mathcal{L}(\bm{\psi}, \mathbf{v}) =
\sum_{i=1}^n 
\frac{\int_{V_i} L(x, y, z) \, dV}{\frac{4}{3}\pi R(v_i)^3}
-
\lambda \cdot \sum_{i=1}^n R(v_i)
-
\begin{cases}
\frac{\lambda}{R(5) - R\left(\sum_{i=1}^n v_i\right)} & \text{if } \sum_{i=1}^n v_i > 0.1 \\
0 & \text{otherwise}
\end{cases}
\]

where radius of the VTA sphere is defined as:

\[
R(v_i) =
\begin{cases}
\sqrt{\dfrac{v_i - 0.1}{0.22}}, & \text{if } v_i > 0.1 \\
0, & \text{otherwise}
\end{cases}
\]

and the sphere's mask is defined voxels within the stimulation sphere's radius:

\[
S(\psi_i, v_i) =
\begin{cases}
1, & \text{if } (x - x_0)^2 + (y - y_0)^2 + (z - z_0)^2 \leq R(v_i)^2 \\
0, & \text{otherwise}
\end{cases}
\]

and the integration of sphere values is empirically the dot product of the sphere and Nifti:

\[
\int_{V_i} L(x, y, z) \, dV \approx \bm{S}(\bm{\psi}_i, v_i)^\top \bm{L}
\]

\textbf{The Gradient Is Found with the Partial Difference Quotient}:

\The partial difference quotient for the gain with respect to \(\mathbf{v}\) at index \(i\) is defined as:

\[
\frac{\partial \mathcal{L}(\bm{\psi}, \mathbf{v})}{\partial v_i} \approx \frac{\mathcal{L}(\bm{\psi}, \mathbf{v} + h\mathbf{e}_i) - \mathcal{L}(\bm{\psi}, \mathbf{v})}{h}
\]

\textbf{Gradient Ascent with Adam Finds the Solution}:

\[
v_i^{(k+1)} = \text{clip}\left[v_i^{(k)} + \alpha \frac{\hat{m}_i^{(k)}}{\sqrt{\hat{v}_i^{(k)}} + \epsilon}\right]
\]

where:
\[
\hat{m}_i^{(k)} = \frac{\beta_1 m_i^{(k)} + (1 - \beta_1) \nabla_{v_i} \mathcal{L}(\bm{\psi}, \mathbf{v})}{1 - \beta_1^{k+1}}, \quad
\hat{v}_i^{(k)} = \frac{\beta_2 v_i^{(k)} + (1 - \beta_2) \left(\nabla_{v_i} \mathcal{L}(\bm{\psi}, \mathbf{v})\right)^2}{1 - \beta_2^{k+1}}
\]

---

\section*{Key Elements: Definitions}

\begin{itemize}
    \item \(\bm{S}(\bm{\psi}_i, v_i)\): Sphere mask for the \(i\)-th contact.
    \item \(\bm{L}\): Landscape values representing magnitudes across a grid of points.
    \item \(r(v_i) = \max\left(0, \sqrt{\frac{v_i - 0.1}{0.22}}\right)\): Radius of the sphere for contact \(v_i\), ensuring \(v_i > 0.1\) to compute nonzero radii.
    \item \(\lambda\): Regularization coefficient penalizing excessive stimulation.
    \item Gradient clipping:
    \[
    \text{clip}\left[v_i\right] = \max(0, \min(v_i, 5))
    \]
    ensures \(v_i\) values remain within the allowable range (0 to 5).
    \item \(\beta_1, \beta_2, \alpha, \epsilon\): Hyperparameters for Adam optimization:
    \begin{itemize}
        \item \(\beta_1\): Momentum decay rate.
        \item \(\beta_2\): Variance decay rate.
        \item \(\alpha\): Learning rate.
        \item \(\epsilon\): Small constant to prevent division by zero.
    \end{itemize}
\end{itemize}

---

\section*{Description}

This function integrates:
\begin{enumerate}
    \item Target Function: Summing contributions of high-value regions across sphere masks, normalized by sphere volume. This forces the contact to stimulate high values, but not expand across the brain.
    \item Penalty Functions:
    \begin{itemize}
        \item Per-contact penalty: Regularizes activation of individual contacts. This suppresses each contact, resulting in it shutting off unless it adds marginal benefit. 
        \item All-contacts penalty: Applies a blocking constraint to prevent total stimulation exceeding the safety limit of 5mA. 
    \end{itemize}
    \item Optimization:
    \begin{itemize}
        \item Gradient ascent with adaptive learning (Adam).
        \item Gradient clipping to stabilize updates.
    \end{itemize}
    \item Stop Conditions: Convergence determined by gradient magnitude (L1 norm) and capped iterations.
\end{enumerate}

---

\section*{Component-Based Breakdown of the Optimization Framework}

Next, we'll go over each component in a more understandable approach. If you've already got it, don't read further. We are going to talk about the mathematical framework for optimizing stimulation parameters \(\bm{\psi}\) and \(\mathbf{v}\). We will expand each part of the loss function into its fundamental components and combine them into the final, simplified formulation.

---

\pagebreaksection

\section*{Component 1: Radius Function \( R(v_i) \)}

The radius \( R(v_i) \) determines the size of the sphere affected by a contact's stimulation amplitude \( v_i \):

\[
R(v_i) =
\begin{cases}
\sqrt{\dfrac{v_i - 0.1}{0.22}}, & \text{if } v_i > 0.1 \\
0, & \text{otherwise}
\end{cases}
\]

\textbf{Purpose:}
\begin{itemize}
    \item Links stimulation amplitude to the physical size of the region of influence.
    \item Prevents stimulation for small or zero amplitudes (\( v_i \leq 0.1 \)).
\end{itemize}

---

\pagebreaksection

\section*{Component 2: Sphere Mask Function \( S(\psi_i, v_i) \)}

The sphere mask \( S(\psi_i, v_i) \) identifies the region of influence around each electrode at position \(\psi_i = (x_0, y_0, z_0)\), compared to all voxels in the Nifti Landscape \(\L(\psi_k) = (x_k, y_k, z_k)\)

\[
S(\psi_i, v_i; \psi_k) =
\begin{cases}
1, & \text{if } (x_k - x_0)^2 + (y_k - y_0)^2 + (z_k - z_0)^2 \leq R(v_i)^2 \quad \forall k \in \{1, \dots, K\} \\
0, & \text{otherwise}
\end{cases}
\]

\text{where:}
\begin{aligned}
    & K && \text{is the total number of voxels in the Nifti landscape,} \\
    & \psi_k = (x_k, y_k, z_k) && \text{are the coordinates of the \(k\)-th voxel,} \\
    & \psi_i = (x_0, y_0, z_0) && \text{are the sphere center coordinates for electrode \(i\), and} \\
    & R(v_i) && \text{is the radius of the sphere defined by the amplitude \(v_i\).}
\end{aligned}

\textbf{Purpose:}
\begin{itemize}
    \item Specifies which points in space are within the sphere defined by \( R(v_i) \).
    \item Encodes spatial information of each sphere’s center \(\psi_i\) and radius \( R(v_i) \).
    \item Promotes density of positive values, avoiding uninhibited expansion of radius.
\end{itemize}

---

\pagebreaksection

\section*{Component 3: Target Function \( T(\bm{\psi}, \mathbf{v}) \)}

The target function measures the effectiveness of stimulation by integrating the empirical landscape \( L(x, y, z) \) within the sphere:

\[
T(\bm{\psi}, \mathbf{v}) =
\sum_{i=1}^n 
\frac{\int_{V_i} L(x, y, z) \, dV}{\frac{4}{3}\pi R(v_i)^3}
=
\sum_{i=1}^n 
\frac{\bm{S}(\psi_i, v_i)^\top \bm{L}}{\frac{4}{3}\pi R(v_i)^3}
\]

\textbf{Purpose:}
\begin{itemize}
    \item Evaluates stimulation effectiveness by weighting the empirical Nifti landscape values \( \bm{L} \) within the sphere.
    \item Normalizes by the volume of the sphere to ensure scale invariance.
    \item The integral is equivalent to a dot product between mask and values \( \bm{S}^\top \bm{L} \).
\end{itemize}

---

\pagebreaksection

\section*{Component 4: Penalty Terms \( P_1 \) and \( P_2 \)}

The penalties \( P_1 \) and \( P_2 \) ensure safe and efficient stimulation:

\[
P_1 = \lambda \cdot \sum_{i=1}^n R(v_i)
\]

\[
P_2 =
\begin{cases}
\frac{\lambda}{R(5) - R\left(\sum_{i=1}^n v_i\right)}, & \text{if } \sum_{i=1}^n v_i > 0.1 \\
0, & \text{otherwise}
\end{cases}
\]

\textbf{Purpose:}
\begin{itemize}
    \item \( P_1 \): Penalizes large spheres to discourage excessive stimulation.
    \item \( P_2 \): Implements a blocking constraint to prevent unsafe total stimulation amplitudes (\( \sum v_i > 5 \)).
\end{itemize}

---

\pagebreaksection

\section*{Component 5: Loss Function \( \mathcal{L}(\bm{\psi}, \mathbf{v}) \)}

The loss function combines the target function \( T \) and penalties \( P_1, P_2 \):

\[
\mathcal{L}(\bm{\psi}, \mathbf{v}) =
T(\bm{\psi}, \mathbf{v}) - P_1(\mathbf{v}) - P_2(\mathbf{v})
\]

Substituting \( T, P_1, P_2 \):

\[
\mathcal{L}(\bm{\psi}, \mathbf{v}) =
\sum_{i=1}^n 
\frac{\bm{S}(\psi_i, v_i)^\top \bm{L}}{\frac{4}{3}\pi R(v_i)^3}
-
\lambda \cdot \sum_{i=1}^n R(v_i)
-
\begin{cases}
\frac{\lambda}{R(5) - R\left(\sum_{i=1}^n v_i\right)}, & \text{if } \sum_{i=1}^n v_i > 0.1 \\
0, & \text{otherwise}
\end{cases}
\]

---

\pagebreaksection

\section*{Component 6: Difference Quotient}

The difference quotient approximates the partial derivative of \(\mathcal{L}\) with respect to \( v_i \):

\[
\frac{\partial \mathcal{L}(\bm{\psi}, \mathbf{v})}{\partial v_i} \approx \frac{\mathcal{L}(\bm{\psi}, \mathbf{v} + h) - \mathcal{L}(\bm{\psi}, \mathbf{v})}{h}
\]

\textbf{Purpose:}
\begin{itemize}
    \item Approximates gradients numerically for optimization.
    \item Evaluates the change in \( \mathcal{L} \) when a small perturbation \( h \) is applied to \( v_i \).
\end{itemize}

---

\pagebreaksection

\section*{Component 7: Gradient Ascent with ADAM Optimization}

The optimization step for \( v_i \) using ADAM is:

\[
v_i^{(k+1)} = \text v_i^{(k)} + \alpha \frac{\hat{m}_i^{(k)}}{\sqrt{\hat{v}_i^{(k)}} + \epsilon}
\]

Where:
\[
\hat{m}_i^{(k)} = \frac{\beta_1 m_i^{(k)} + (1 - \beta_1) \nabla_{v_i} \mathcal{L}(\bm{\psi}, \mathbf{v})}{1 - \beta_1^{k+1}}, \quad
\hat{v}_i^{(k)} = \frac{\beta_2 v_i^{(k)} + (1 - \beta_2) \left(\nabla_{v_i} \mathcal{L}(\bm{\psi}, \mathbf{v})\right)^2}{1 - \beta_2^{k+1}}
\]

\textbf{Purpose:}
\begin{itemize}
    \item Dynamically adjusts step sizes using moment estimates \( m_i, v_i \).
    \item Prevents destabilizing oscillation, which our gradients are sensitive to.
    \item Prevents gradient explosion or vanishing.
\end{itemize}

---

\pagebreaksection

\section*{Summary: The Simplified Loss Function}

The simplified formulation is:

\[
\mathcal{L}(\bm{\psi}, \mathbf{v}) =
\sum_{i=1}^n 
\frac{\bm{S}(\psi_i, v_i)^\top \bm{L}}{\frac{4}{3}\pi R(v_i)^3}
-
\lambda \cdot \sum_{i=1}^n R(v_i)
-
\begin{cases}
\frac{\lambda}{R(5) - R\left(\sum_{i=1}^n v_i\right)}, & \text{if } \sum_{i=1}^n v_i > 0.1 \\
0, & \text{otherwise}
\end{cases}
\]

This equation integrates all components \( R, S, T, P_1, P_2 \) into a single mathematical representation.

\section*{Optimization Loop for Loss Function \(\mathcal{L}(\bm{\psi}, \mathbf{v})\)}

The optimization process iteratively updates the stimulation amplitudes \(\mathbf{v}\) using gradient ascent with ADAM optimization. The steps are as follows:

\subsection*{Step 1: Compute the Gradient Vector}
The gradient of the loss function with respect to each amplitude \(v_i\) is computed using the partial difference quotient:
\[
\frac{\partial \mathcal{L}(\bm{\psi}, \mathbf{v})}{\partial v_i} \approx \frac{\mathcal{L}(\bm{\psi}, \mathbf{v} + h) - \mathcal{L}(\bm{\psi}, \mathbf{v})}{h}
\]

\subsection*{Step 2: Apply Gradient Clipping}
The gradient vector \(\nabla_{\mathbf{v}} \mathcal{L}\) is clipped to prevent instability:
\[
\nabla_{\text{clipped}} = \text{clip}\left(\nabla_{\mathbf{v}}, -\lambda_{\text{max}}, \lambda_{\text{max}}\right)
\]

\subsection*{Step 3: Perform ADAM Optimization}
ADAM updates the amplitudes based on moment estimates:
\[
v_i^{(k+1)} = \text v_i^{(k)} + \alpha \frac{\hat{m}_i^{(k)}}{\sqrt{\hat{v}_i^{(k)}} + \epsilon}
\]
where:
\[
\hat{m}_i^{(k)} = \frac{\beta_1 m_i^{(k)} + (1 - \beta_1) \nabla_{\text{clipped}}}{1 - \beta_1^{k+1}}, \quad
\hat{v}_i^{(k)} = \frac{\beta_2 v_i^{(k)} + (1 - \beta_2) (\nabla_{\text{clipped}})^2}{1 - \beta_2^{k+1}}
\]

\subsection*{Step 4: Enforce Constraints on \(\mathbf{v}\)}
No negative amperages allowed, so apply constraints:
\[
v_i \geq 0 \quad
\]

\subsection*{Step 5: Check Stopping Conditions}
Stop the optimization if:
\begin{itemize}
    \item The gradient's L1 norm is below a threshold:
    \[
    \|\nabla_{\mathbf{v}} \mathcal{L}\|_1 < \tau
    \]
    \item The maximum number of iterations is reached (iterations = 50).
\end{itemize}

\newpage

\subsection*{Overall Loop}
Repeat the following steps until convergence or stopping criteria are met:
\[
\textbf{Loop:} 
\begin{cases}
\text{1. Compute Gradient} \\
\text{2. Apply Gradient Clipping} \\
\text{3. Perform ADAM Update} \\
\text{4. Enforce Constraints} \\
\text{5. Check Stopping Conditions}
\end{cases}
\]

\subsection*{Step 6: Final Step}
Once the loop has concluded, return the optimal amperages at each contact:
\[
\mathbf{v}_{\text{optimal}} = \{v_1, v_2, \dots, v_n\}
\]

\begin{flushright}
\textit{And that's all there is to it.} \\
-- Calvin
\end{flushright}

\end{document}

